Manuscript results: Forecasting influenza using ARIMA and ARIMAX

### 2025-06-03 manuscript_results.Rmd
update

```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
library("ggplot2")
library("sf")
library("forcats")
library("ggplot2")
library("sf")
library("dplyr")
library("scales")  # For `breaks_extended`
library("ggplot2")
library("broom")
library("fields")
library("ggpubr")
library("patchwork")
library("ggpattern")

@@

custom_colors <- c(
  "AUTO_AR_LB" = colors_ar[1], "ES27_AR_LB" = colors_ar[1], "ES64_AR_LB" = colors_ar[1],
  "AUTO_EPI_LB" = colors_epi[1], "ES27_EPI_LB" = colors_epi[1], "ES64_EPI_LB" = colors_epi[1],
  "AUTO_TMP_LB" = colors_tmp[1], "ES27_TMP_LB" = colors_tmp[1], "ES64_TMP_LB" = colors_tmp[1],
  "AUTO_ADJ_LB" = colors_adj[1], "ES27_ADJ_LB" = colors_adj[1], "ES64_ADJ_LB" = colors_adj[1],
  "AUTO_AVG_LB" = colors_avg[1], "ES27_AVG_LB" = colors_avg[1], "ES64_AVG_LB" = colors_avg[1],

@@

plot4

@@

library(ggpattern)
library(ggplot2)
library(dplyr)

# Create the patterned plot
combined_plot <- ggplot(all_weeks, aes(x = Best_Result, 
                                       fill = Best_Result,
                                       pattern = Best_Result)) +
  geom_bar_pattern(
    pattern_fill  = "black",
    pattern_size  = 0.1,
    pattern_angle = 45,  # set globally
    colour        = NA
  ) +
  scale_fill_manual(values = custom_colors) +
  scale_pattern_manual(values = custom_patterns) +  # define this before
  labs(title = "Best model based on the lowest mean WIS",
       subtitle = "Results for the 48 states on the contiguous U.S. by target week",
       x = "", y = "Number of states",
       fill = "Best model",
       pattern = "Best model") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 16, face = "bold"),legend.position = "none") +
  facet_grid(Week_Ahead ~ Model_Type, scales = "free_x") +
  guides(pattern_angle = "none")  # just in case

# Save and display the plot
ggsave("Fig4.jpg", combined_plot, width = 7, height = 6)
print(combined_plot)
```

@@

```{r}
calculate_percentage_of_improvement <- function(data) {
  return(data.frame(
    AUTO_AR = (data$AUTO_AR / data$AUTO_AR),

@@

@@

# Save to file
ggsave("Fig6.jpg", combined_plot, width =12, height =7)

@@

# Combine the plots vertically
reg2_scatter <- reg2 | combined_regressions 
reg2_scatter
ggsave(reg2_scatter, height = 6, width = 12, filename="Fig7.jpg")  
#ggsave(reg2, filename="reg2.jpg",height = 8, width = 10)
```

@@

# Save to file
ggsave("Fig8.jpg", combined_plot, width =12, height =7)
```

@@

# Plot
reg2<-ggplot(regression_models, aes(x = Week_Ahead, y = independent_variable, fill = significance)) +  
  geom_tile(color = "black") +  # Add black borders to squares
  geom_text(aes(label = round(r_squared, 3)), color = "black", size = 6) +  # Text inside boxes
  scale_fill_manual(values = c("Significant" = "cyan", "Not Significant" = "gray")) +  
  labs(title = "Regression models R²",
       subtitle = "Mean WIS differences (%) as dependent variable",
       x = "Target Week",
       y = "Independent Variables",
       fill = "Significance (p < 0.05)") +    
  theme_minimal() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size = 16, face = "bold"),
                  plot.subtitle = element_text(size = 14))

@@

reg2_scatter <- reg2 | combined_regressions 
reg2_scatter
ggsave(reg2_scatter, height = 6, width = 12, filename="Fig9.jpg")  
#ggsave(reg2, filename="Fig9.jpg",height = 6, width = 8)
```

@@

# Plot the difference using ggplot2
model_improv<-ggplot(summed_data_all, aes(x = Julian_date, y = (Difference), color = Week)) +
  geom_point(size = 2.5) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed", size = 0.8) +
  theme_minimal() +
  labs(
    title = "Weekly mean WIS differences across all states",
    subtitle = "AUTO_AVG_LB compared to baseline",
    x = "",
    y = "Mean difference"
  ) +
  scale_y_continuous(
    limits = c(-100, 100),
    breaks = seq(-100, 100, by = 20)
  ) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %y"
  ) +
  scale_color_manual(
    values = c(
    "Week 1" = "#00BFC4",  # Bright cyan
    "Week 2" = "#F8766D",  # Bright coral red
    "Week 3" = "#7CAE00",  # Lime green
    "Week 4" = "#C77CFF"   # Bright purple
  )
  ) +
  theme(
    plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

@@

hospital_plot <-ggplot(total_hospitalizations_by_week, aes(x = target_end_date, y = mean_cases)) +
  geom_bar(stat = "identity",fill = "steelblue") +
  labs(title = "Total influenza hospitalizations on the 48 states",
       x = "",
       y = "Influenza hospitalizations") +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %y"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1))

@@

# Combine the plots vertically
combined_plot <- hospital_plot / model_improv

@@

# Save the combined plot
ggsave("Fig10.jpg", combined_plot, width = 7, height = 7, dpi = 600)

>>>>

```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
library("ggplot2")
library("sf")
library("forcats")
library("ggplot2")
library("sf")
library("dplyr")
library("scales")  # For `breaks_extended`
library("ggplot2")
library("broom")
library("fields")
library("ggpubr")
library("patchwork")
library("ggpattern")
library("cowplot")
```

@@

```{r}
custom_colors <- c(
  "AUTO_AR_LB" = colors_ar[1], "ES27_AR_LB" = colors_ar[1], "ES64_AR_LB" = colors_ar[1],
  "AUTO_EPI_LB" = colors_epi[1], "ES27_EPI_LB" = colors_epi[1], "ES64_EPI_LB" = colors_epi[1],
  "AUTO_TMP_LB" = colors_tmp[1], "ES27_TMP_LB" = colors_tmp[1], "ES64_TMP_LB" = colors_tmp[1],
  "AUTO_ADJ_LB" = colors_adj[1], "ES27_ADJ_LB" = colors_adj[1], "ES64_ADJ_LB" = colors_adj[1],
  "AUTO_AVG_LB" = colors_avg[1], "ES27_AVG_LB" = colors_avg[1], "ES64_AVG_LB" = colors_avg[1],

@@

plot4
```

# Create the patterned plot
combined_plot <- ggplot(all_weeks, aes(x = Best_Result, 
                                       fill = Best_Result,
                                       pattern = Best_Result)) +
  geom_bar_pattern(
    pattern_fill  = "black",
    pattern_size  = 0.1,
    pattern_angle = 45,  # set globally
    colour        = NA
  ) +
  scale_fill_manual(values = custom_colors) +
  scale_pattern_manual(values = custom_patterns) +  # define this before
  labs(title = "Best model based on the lowest mean WIS",
       subtitle = "Results for the 48 states on the contiguous U.S. by target week",
       x = "", y = "Number of states",
       fill = "Best model",
       pattern = "Best model") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 16, face = "bold"),legend.position = "left") +
  facet_grid(Week_Ahead ~ Model_Type, scales = "free_x") #+
 # guides(pattern_angle = "none")  # just in case

# Save and display the plot
#ggsave("Fig4.jpg", combined_plot, width = 7, height = 6)
print(combined_plot)

@@

# Define the custom patterns
custom_patterns <- c("AUTO" = "circle", "ES27" = "stripe", "ES64" = "crosshatch")
custom_fills <- c("AUTO" = "white", "ES27" = "white", "ES64" = "white")  # or any colors you like

# Create the custom legend plot
legend_data <- data.frame(
  Model_Type = c("AUTO", "ES27", "ES64", "ES27", "ES64"),
  Model_Family = c("AR","TMP", "EPI", "ADJ","AVG"),
  x = 1:5,
  y = 1
)

# Define patterns for model types
custom_patterns <- c("AUTO" = "circle", "ES27" = "stripe", "ES64" = "crosshatch")

# Define fill colors for approaches
custom_fills <- c("AR" = colors_ar[1], "TMP"=colors_tmp[1], "EPI" = colors_epi[1], "ADJ" = colors_adj[1], "AVG" = colors_avg[1])

# Custom legend plot
custom_legend <- ggplot(legend_data, aes(x = factor(x), y = y, 
                                         pattern = Model_Type, fill = Model_Family)) +
  geom_bar_pattern(stat = "identity",
                   pattern_fill = "black",
                   pattern_density = 0.3,
                   pattern_spacing = 0.05,
                   pattern_size = 0.15,
                   colour = "black",
                   width = 0.7) +
  scale_fill_manual(values = custom_fills, name = "Model Family") +
  scale_pattern_manual(values = custom_patterns, name = "Model Type") +
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10)
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(pattern = "none", pattern_fill = NA, colour = NA)
    ),
    pattern = guide_legend(
      override.aes = list(fill = "white", pattern_fill = "black")
    )
  )
custom_legend
```

# Remove legends from combined_plot
combined_plot_clean <- combined_plot + theme(legend.position = "none")

# Extract only the legend from custom_legend

legend_only <- cowplot::get_legend(custom_legend)

# Combine your main plot with the custom legend
final_plot <- combined_plot_clean + 
  patchwork::plot_spacer() + 
  patchwork::wrap_elements(legend_only) + 
  plot_layout(widths = c(4, 0.05, 0.4))  # Adjust widths as needed

# Display the result
final_plot

ggsave("Fig4.jpg", final_plot, width = 10, height = 6)
```

@@

```{r}
calculate_percentage_of_improvement <- function(data) {
  return(data.frame(
    AUTO_AR = (((data$AUTO_AR / data$AUTO_AR)-1) * 100),

@@

new_long_data <- summary_impr %>%
  mutate(
    LogBack = grepl("_LB$", Model),
    Family = sub("_.*", "", Model),  # take everything before the first underscore
    Method = sub("^.*?_(AR|ADJ|EPI|AVG|TMP)(_LB)?$", "\\1", Model)
  )

new_long_data <- new_long_data %>%
  mutate(Model = ifelse(Model == "AUTO_AR", "AUTO_AAR", Model))

model <- lm(m ~ Model, data = new_long_data)
summary(model)

model
new_long_data$Model
# Get tidy summary of coefficients
tidy_model <- broom::tidy(model)

# Filter out intercept for clarity if you want
tidy_model_no_intercept <- tidy_model %>% filter(term != "(Intercept)")

# Plot coefficients with 95% confidence intervals
tidy_model_no_intercept <- tidy_model_no_intercept %>%
  mutate(term = forcats::fct_reorder(term, estimate, .desc = TRUE))

# Plot
ggplot(tidy_model_no_intercept, aes(x = term, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = estimate - std.error * 1.96,
                    ymax = estimate + std.error * 1.96),
                width = 0.2) +
  coord_flip() +  # flip for better readability
  labs(
    title = "Effect Size of Model Types on outcome",
    subtitle =  "Estimates relative to baseline (intercept)",
    y = "Estimate",
    x = ""
  ) +
  theme_minimal()

new_long_data <- summary_impr %>%
  mutate(
    LogBack = grepl("_LB$", Model),
    Type = sub("_.*", "", Model),  # take everything before the first underscore
    Family = sub("^.*?_(AR|ADJ|EPI|AVG|TMP)(_LB)?$", "\\1", Model)
  )

new_long_data <- new_long_data %>%
  mutate(Family = ifelse(Family == "AR", "AAR", Family))

model <- lm(m ~ Type + Family + LogBack, data = new_long_data)
summary(model)

model
new_long_data$Model
# Get tidy summary of coefficients
tidy_model <- broom::tidy(model)

# Filter out intercept for clarity if you want
tidy_model_no_intercept <- tidy_model %>% filter(term != "(Intercept)")

# Plot coefficients with 95% confidence intervals
tidy_model_no_intercept <- tidy_model_no_intercept %>%
  mutate(term = forcats::fct_reorder(term, estimate, .desc = TRUE))

# Plot
effect_size<- ggplot(tidy_model_no_intercept, aes(x = term, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = estimate - std.error * 1.96,
                    ymax = estimate + std.error * 1.96),
                width = 0.2) +
  coord_flip() +  # flip for better readability
  labs(
    title = "Effect size of model type, family and transformation on WIS difference",
    subtitle =  "WIS Difference relative to the AUTO type, ARIMA family and LogBack False (intercept)",
    y = "WIS Difference",
    x = ""
  ) +
  theme_minimal() + theme(plot.title = element_text(size = 16, face = "bold"),
                  plot.subtitle = element_text(size = 14))

effect_size

ggsave("Fig6.jpg", effect_size, width =10, height =5)

@@

# Save to file
ggsave("Fig7.jpg", combined_plot, width =12, height =7)

@@

# Combine the plots vertically
reg2_scatter <- reg2 | combined_regressions 
reg2_scatter
ggsave(reg2_scatter, height = 6, width = 12, filename="Fig8.jpg")  
#ggsave(reg2, filename="reg2.jpg",height = 8, width = 10)
```

@@

# Save to file
ggsave("Fig9.jpg", combined_plot, width =12, height =7)
```

@@

# Plot
reg2<-ggplot(regression_models, aes(x = Week_Ahead, y = independent_variable, fill = significance)) +  
  geom_tile(color = "black") +  # Add black borders to squares
  geom_text(aes(label = round(r_squared, 3)), color = "black", size = 6) +  # Text inside boxes
  scale_fill_manual(values = c("Significant" = "cyan", "Not Significant" = "gray")) +  
  labs(title = "Regression models R²",
       subtitle = "Mean WIS differences as dependent variable",
       x = "Target Week",
       y = "Independent Variables",
       fill = "Significance (p < 0.05)") +    
  theme_minimal() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        axis.text.y = element_text(size = 10),
        plot.title = element_text(size = 16, face = "bold"),
                  plot.subtitle = element_text(size = 14))

@@

reg2_scatter <- reg2 | combined_regressions 
reg2_scatter
ggsave(reg2_scatter, height = 6, width = 12, filename="Fig10.jpg")  
#ggsave(reg2, filename="Fig9.jpg",height = 6, width = 8)
```

@@

# Plot the difference using ggplot2
model_improv<-ggplot(summed_data_all, aes(x = Julian_date, y = (Difference), color = Week)) +
  geom_point(size = 2.5) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed", size = 0.8) +
  theme_minimal() +
  labs(
    title = "A) Weekly mean WIS differences across all states",
    subtitle = "AUTO_AVG_LB compared to baseline",
    x = "",
    y = "Mean difference"
  ) +
  scale_y_continuous(
    limits = c(-100, 100),
    breaks = seq(-100, 100, by = 20)
  ) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %y"
  ) +
scale_color_manual(
  values = c(
    "Week 1" = "#4575B4",  # Blue
    "Week 2" = "#91BFDB",  # Light blue
    "Week 3" = "#E89C9C",  # Light red
    "Week 4" = "#D73027"   # Red
  )
) +
  theme(
    plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

@@

hospital_plot <-ggplot(total_hospitalizations_by_week, aes(x = target_end_date, y = mean_cases)) +
  geom_bar(stat = "identity",fill = "steelblue") +
  labs(title = "B) Total influenza hospitalizations on the 48 states",
       x = "",
       y = "Influenza hospitalizations") +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %y"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1))

@@

# Combine the plots vertically
combined_plot <-  model_improv /hospital_plot

@@

# Save the combined plot
ggsave("Fig11.jpg", combined_plot, width = 7, height = 7, dpi = 600)
